Large and Infinitary Quotient Inductive-Inductive Types

### 2020-01-10 draft/paper.tex

Cartmell \cite{gat}. Displayed categories Ahrens \cite{displayedcats}.

Paolo paper \cite{gabe}. Finitary case \cite{kaposi2019constructing}.

Direct construction of concrete QIIT: Streicher
\cite{streicher2012semantics}, Brunerie \cite{brunerie}, Ambroise
\cite{induction-is-enough}.

Higher inductive types: HIT descriptions \cite{niels,hiit}, HITs in
cubical type theories\cite{cubicalhits,10.1145/3290314}, semantics for
subclasses of HITs
\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}
Construction of example IITs without UIP: Jasper \cite{jasper}.

\section{Conclusion}
\label{sec:conclusion}

>>>>

Cartmell \cite{gat} defines generalised algebraic theories (GATs)
using type theoretic syntax. Compared to our QII signatures, his
signatures support sort equations but do not cover infinitary
constructors or recursive equations. A way to encode sort equations in
our signatures is using isomorphisms instead of equalities. It seems
that infinitary constructors and sort equations are not compatible in
a setting with UIP. In contrast to our algebraic definition,
Cartmell's signatures are given by presyntax, named variables and
typing relations. He also does not construct categories of algebras
explicitly, but is concerned with showing that the classifying
category for GATs is that of contextual categories.

A more semantic approach to QIITs is given by Altenkirch et al.\
\cite{gabe}. They generalise the usual functorial semantics of
inductive types to QIITs by considering towers of functors and
building complete categories of algebras from such towers. Their
notion of signature does not enforce strict positivity, hence
describes a larger class of QII signatures. The equivalence of
initiality and induction is showed, but initial algebras are not
constructed.

Kaposi et al's work \cite{kaposi2019constructing} is the most direct
precursor of our work. They don't consider infinitary constructors or
constructors with recursive equations which make their semantics much
simpler.

Streicher's initiality proof \cite{streicher2012semantics} is a direct
construction of a concrete QIIT: that of the syntax of type theory. He
uses preterms, typing relations and quotients to construct the initial
model of type theory in a set-theoretic metalanguage. A recent
formalisation of the same result for a larger type theory was done by
Brunerie et al.\ \cite{brunerie} in Agda with UIP, function
extensionality and quotients. Lafont et al.\
\cite{induction-is-enough} construct all finitary inductive-inductive
types in Agda using only UIP and function extensionality. Examples of
inductive-inductive types are constructed without using UIP by Hugunin
\cite{jasper}.

Higher inductive types (HITs) are generalisations of QIITs in settings
with proof-relevant identity types, in fact, they were introduced
before QIITs \cite{HoTTbook}. \cite{niels} is a description of a
subset of HITs. \cite{hiit} describes a syntax for higher
inductive-inductive types using a theory of signatures similar to ours,
however does not construct categories of algebras or show
initiality. Semantics for different subclasses of HITs are given by
\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
type theories were shown to support HITs in a computational way
\cite{niels,hiit}.

Our notion of displayed CwF is an extension of displayed category
\cite{displayedcats} however in a setting with UIP.

\section{Conclusions and further work}
>>>>>>> eba72f32c6607b337c6315a124afe03f6f3ad92d
\label{sec:conclusion}

Existence of the theory of signatures:

\cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
constructible in type theory without choice, so the existence of the
theory of signatures is a non-conservative extension of extensional
type theory. The setoid model supports an example QIT which branches
over an arbitrary setoid, and this suggests that it might support the
theory of signatures as well.
